# Methods

```{r}
#| label: methods-setup
#| include: false

library(knitr)

# Project root
project_root <- normalizePath(
  file.path(getwd(), ".."),
  winslash = "/",
  mustWork = TRUE
)

# Manuscript figure directory
fig_dir <- file.path(project_root, "figs")

# Helper for manuscript figures
fig_path <- function(filename) {
  file.path(fig_dir, filename)
}

# Fail clearly if a required figure is missing
require_file <- function(path) {
  if (!file.exists(path)) {
    stop("Required manuscript file not found: ", path)
  }
  invisible(path)
}

# Include an already-generated manuscript figure
include_manuscript_figure <- function(filename) {
  path <- fig_path(filename)
  require_file(path)

  knitr::include_graphics(
    path,
    rel_path = FALSE
  )
}

```

## Study area and acoustic surveys

Acoustic surveys were conducted at eight sites in Plumas County, California, in the northern Sierra Nevada, spanning an elevational gradient of 581–1,846 m. Most sites were near perennial water sources, including reservoirs (Bucks Lake, Lower Bucks Lake, Grizzly Forebay, and Three Lakes), tributary streams, and the North Fork Feather River; one site was situated in an upland meadow adjacent to forest (@fig-map). Surveys were conducted from April through October 2015, with detectors programmed to record from one hour before sunset to one hour after sunrise.

```{r}
#| label: fig-map
#| fig-cap: "Locations of eight passive acoustic monitoring sites in Plumas County, California, USA, in the northern Sierra Nevada. Detectors were distributed across an elevational gradient and included sites associated with Bucks Lake, Lower Bucks Lake, Grizzly Forebay, Three Lakes, and surrounding stream and upland habitats. The inset shows the location of the study area within northern California."
#| echo: false
#| out-width: "85%"
#| fig-align: center

include_manuscript_figure("Figure1_map.PNG")
```

Each site was equipped with an Anabat II ultrasonic detector and CF-ZCAIM data storage unit (Titley Electronics, Ballina, NSW, Australia) housed in a weatherproof enclosure mounted at least 1 m above the ground. Detectors were fitted with a microphone extension cable, weather shroud, and flat acoustic reflector. Units were powered by batteries recharged with 5-W solar panels and miniature charge controllers (EME Systems, Berkeley, California, USA). Detectors recorded nightly temperature and relative humidity. Lunar phase was obtained from the U.S. Naval Observatory and represented on a scale from 0 (new moon) to 1 (full moon). This variable indexes position in the lunar cycle rather than realized nocturnal illumination at individual sites.

## Acoustic classification and activity index

Echolocation calls were classified using Kaleidoscope Pro (Wildlife Acoustics Inc., version 5.1.8) with the Bats of North America 5.1.0 classifier at the "More Accurate (Conservative)" setting. Ten percent of classified calls were manually vetted.

Bat activity was quantified using a modified Acoustic Activity Index [@millerMethodDeterminingRelative2001] based on 5-min sampling intervals. For each site-night, we counted the number of 5-min intervals containing at least one accepted bat detection. The response therefore represents the number of intervals in which bats were acoustically detected rather than the number of individual calls or bats.

Because environmental measurements could be available on nights when acoustic sampling failed, we reconstructed detector effort independently of bat detections. Site-nights lacking evidence of usable acoustic recording were treated as unsampled and excluded from analysis; they were not assigned zero bat activity. Zero-activity observations were retained only when usable acoustic recordings demonstrated that the detector operated but no accepted bat detections occurred. This distinction prevented intermittent equipment failure from being represented as ecological absence in the seasonal and environmental response models.

For environmental analyses, nightly mean temperature and relative humidity were standardized to a mean of zero and standard deviation of one. Day of year and lunar phase were retained on their original scales for model fitting.

## Causal structure and estimands

We used the causal structure in @fig-dag to distinguish seasonal variation from conditional associations with nightly environmental conditions. Day of year was represented as a common cause of temperature, relative humidity, and bat acoustic activity, and temperature was additionally allowed to influence relative humidity. Temperature, relative humidity, and lunar phase were each represented as predictors of acoustic activity. The DAG was used to define the quantities estimated by the seasonal and environmental models rather than to infer causal effects from the observational data alone.

```{r}
#| label: fig-dag
#| fig-cap: "Directed acyclic graph representing the hypothesized relationships among day of year, nightly temperature, relative humidity, lunar phase, and bat acoustic activity. Arrows represent assumed causal relationships used to guide model specification and interpretation. Day of year represents seasonal structure in activity and environmental conditions not otherwise represented by the measured nightly covariates. Temperature was hypothesized to influence relative humidity in addition to bat activity; relative humidity and lunar phase were also hypothesized to influence activity. The graph was used to define the total seasonal and conditional environmental estimands."
#| echo: false
#| out-width: "85%"
#| fig-align: center

include_manuscript_figure("Figure2_DAG.PNG")
```

The total seasonal model estimated variation in acoustic activity across day of year without conditioning on nightly temperature, relative humidity, or lunar phase. Because environmental conditions are themselves seasonally structured, this model represents the seasonal activity pattern expressed under the environmental conditions occurring during the study.

The conditional environmental model included day of year, nightly temperature, relative humidity, and lunar phase simultaneously. Environmental response curves therefore represent associations with each focal predictor conditional on the remaining modeled predictors and site-specific seasonal phenology. The day-of-year component of this model represents seasonal structure remaining after accounting for the measured nightly environmental conditions. Comparing the total and conditional seasonal trajectories allowed us to distinguish the observed seasonal pattern from seasonal structure not associated with the measured nightly covariates.

## Characteristic echolocation-frequency groups

To test whether environmental responses varied with characteristic echolocation frequency, we classified each species into one of three frequency groups: low (\<25 kHz), mid (25–40 kHz), and high (\>40 kHz). Characteristic frequency ($F_c$) was defined as the frequency at the lowest, flattest portion of the descending frequency-modulated sweep of an echolocation call, where call bandwidth is narrowest and the call typically provides its diagnostic signal for species identification [@fentonRecognitionSpeciesInsectivorous1981; @ofarrellQualitativeIdentificationFreeFlying1999]. Group boundaries were based on species accounts in the Western Bat Working Group reference and the Kaleidoscope Pro classifier reference-call library and were chosen to represent broad frequency ranges over which atmospheric attenuation of ultrasound is expected to differ in its sensitivity to environmental conditions [@griffinImportanceAtmosphericAttenuation1971; @lawrenceMeasurementsAtmosphericAttenuation1982].

The low-frequency group contained *Lasiurus cinereus*, *Eumops perotis*, and *Euderma maculatum*. The mid-frequency group contained *Eptesicus fuscus*, *Lasionycteris noctivagans*, *Tadarida brasiliensis*, *Antrozous pallidus*, *Myotis evotis*, *Myotis thysanodes*, and *Corynorhinus townsendii*. The high-frequency group contained *Myotis yumanensis*, *M. lucifugus*, *M. californicus*, *M. ciliolabrum*, *M. volans*, *Parastrellus hesperus*, and *Lasiurus frantzii*. Full species-level information, characteristic frequencies, and detection totals are provided in Table S4.

For each frequency group, nightly activity at each site was calculated as the number of 5-min intervals containing at least one detection from any species assigned to that group. Frequency-group activity was analyzed using the same negative-binomial likelihood, site-specific seasonal structure, environmental predictors, and model-assessment framework used for aggregate activity.

## Bayesian generalized additive models

We modeled nightly acoustic activity using Bayesian generalized additive models (GAMs) with a negative-binomial distribution and log link. The negative-binomial likelihood accommodated substantial among-night variation in the number of occupied 5-min intervals.

We represented seasonal activity using a population-level smooth of day of year together with a factor-smooth interaction between day of year and site. The factor smooth estimated penalized site-specific departures from the overall seasonal trajectory, allowing the timing and magnitude of phenology to differ among sites while sharing information across sites.

The total seasonal model was

$$
n_{it} \sim \mathrm{NegBinomial}(\mu_{it}, \phi),
$$

$$
\log(\mu_{it}) =
\beta_0 + f(\mathrm{DOY}_{it}) + f_{\mathrm{site}}(\mathrm{DOY}_{it}),
$$

where $n_{it}$ is the number of 5-min intervals containing at least one bat detection at site $i$ on night $t$, $f(\mathrm{DOY})$ represents the overall seasonal trajectory, and $f_{\mathrm{site}}(\mathrm{DOY})$ represents penalized site-specific departures from that trajectory.

The conditional environmental model was

$$
\log(\mu_{it}) =
\beta_0 +
f(\mathrm{DOY}_{it}) +
f_{\mathrm{site}}(\mathrm{DOY}_{it}) +
f(T_{it}) +
f(RH_{it}) +
f(L_{it}),
$$

where $T$ is standardized nightly mean temperature, $RH$ is standardized nightly mean relative humidity, and $L$ is lunar phase.

Smooth terms were estimated using the default thin-plate regression spline basis in `brms`/`mgcv`, and site-specific seasonal trajectories were represented using a factor-smooth basis (`bs = "fs"`). Because the factor smooth is fully penalized and represents site-level differences in the seasonal trajectory, we did not include an additional site random intercept.

## Priors, computation, and model assessment

We used a weakly informative Normal(0, 10) prior for the intercept; smoothing parameters and the negative-binomial shape parameter used the regularizing defaults implemented by `brms` [@brms]. Models were estimated using Hamiltonian Monte Carlo, with four chains of 4,000 iterations, including 1,000 warmup iterations per chain, yielding 12,000 post-warmup draws. Sampling used `adapt_delta = 0.99` and a maximum tree depth of 12.

We assessed convergence and sampling quality using $\hat{R}$, bulk and tail effective sample sizes, divergent transitions, maximum-tree-depth warnings, and energy diagnostics. Model adequacy was evaluated using posterior predictive checks of the response distribution and site-level variation, including the mean, standard deviation, and frequency of zero-activity nights. Because sampling was interrupted at some sites by detector failures, residual temporal dependence was evaluated using calendar-day rather than row-order lags so that gaps in sampling were not treated as consecutive observations. Additional temporal structure was considered only when residual dependence indicated lack of fit relevant to the ecological estimands.

Posterior summaries are reported as medians with 89% credible intervals. Environmental response curves were generated by varying the focal environmental predictor across its observed range, averaging predictions over the observed distribution of sampled days of year and weighting sites equally, while holding the remaining environmental predictors at their reference values. Environmental contrasts were calculated between the 10th and 90th percentiles of observed temperature and relative humidity and between new moon (0) and full moon (1) for lunar phase. Seasonal predictions were averaged equally across sites. Predictions from the total seasonal model represent the overall seasonal pattern, whereas predictions from the conditional environmental model represent seasonal structure after accounting for nightly temperature, relative humidity, and lunar phase.

All analyses were conducted in R [@R]. Bayesian generalized additive
models were fit with `brms` [@brms] through `cmdstanr` [@cmdstanr],
with spline bases implemented in `mgcv` [@mgcv]. Posterior summaries
and predictions were generated using `posterior` [@posterior] and
`tidybayes` [@tidybayes].
