# Results

```{r}
#| label: results-setup
#| include: false

library(knitr)

# Project root
project_root <- normalizePath(
  file.path(getwd(), ".."),
  winslash = "/",
  mustWork = TRUE
)

# Manuscript figure directory
fig_dir <- file.path(project_root, "figs")

# Helper for manuscript figures
fig_path <- function(filename) {
  file.path(fig_dir, filename)
}

# Fail clearly if a required figure is missing
require_file <- function(path) {
  if (!file.exists(path)) {
    stop("Required manuscript file not found: ", path)
  }
  invisible(path)
}

# Include an already-generated manuscript figure
include_manuscript_figure <- function(filename) {
  path <- fig_path(filename)
  require_file(path)

  knitr::include_graphics(
    path,
    rel_path = FALSE
  )
}

```

```{=html}
<!--
PURPOSE OF THIS FILE

This Results file is deliberately separated into:
1. results that are already supported by the corrected data and final aggregate models;
2. clearly marked placeholders for frequency-group results that should be populated
   only after the final diagnostic/production fits have been accepted.

The prose is written for a Functional Ecology-style manuscript: lead with the
biological pattern, support it with model output, and keep model-diagnostic detail
brief in the main text while directing fuller diagnostics to the Supplementary
Material.

Do not restore numerical results from an older model merely because they appeared
in a previous manuscript draft. Recompute figure-derived values from the final
models wherever possible.
-->
```

## Sampling and acoustic activity

After accounting for nights without usable acoustic recordings, the environmental-model dataset contained 1,056 site-nights from eight sites. Sampling extended from late April through mid-October 2015. Sampling effort was uneven among sites because of detector failures and other periods without usable acoustic data, with the number of retained nights ranging from 47 at BUL18 to 174 at several sites.

Across these 1,056 site-nights, aggregate bat activity totaled 103,681 5-min detection intervals. The species-level dataset contained 17 species, and the frequency-group datasets were constructed for the same 1,056 site-nights, with 7,078, 27,602, and 69,001 detection intervals assigned to the low-, mid-, and high-frequency groups, respectively.

```{=html}
<!--
TODO TABLE 1:
Generate descriptive environmental ranges from the corrected 1,056-site-night
dataset, not from the original uncorrected dataset.

Suggested columns:
Variable | Median | Lower range | Upper range

Likely variables:
- Day of year / sampling dates
- Nightly mean temperature (°C)
- Relative humidity (%)
- Lunar phase (0–1)

Do not hard-code the old manuscript ranges until they have been checked against
agg_data_corrected / the data used by m_agg_env.
-->
```

## Model development and adequacy

Model development materially improved representation of the observed ecological and sampling structure. Initial Poisson models showed strong residual temporal dependence, with lag-1 residual correlation of approximately 0.80 and correlations remaining above 0.50 through a 14-day lag. Replacing the Poisson likelihood with a negative-binomial likelihood accounted for substantial extra-Poisson variation but did not eliminate the temporal structure.

Allowing seasonal phenology to vary among sites using a penalized factor-smooth interaction produced a much larger improvement. In the final aggregate conditional model, lag-1 residual correlation was 0.379, declined to 0.208 at lag 2 and 0.168 at lag 3, was near zero by lag 4, and became weakly negative at longer lags. This pattern indicated that much of the apparent serial dependence in simpler models reflected unmodeled differences in seasonal phenology among sites rather than a persistent autoregressive observation process.

The final aggregate conditional model also showed satisfactory MCMC behavior. All reported $\hat{R}$ values were 1.00, with large bulk and tail effective sample sizes. Seven of 12,000 post-warmup transitions reached the maximum tree depth of 12. Posterior predictive checks reproduced the broad response distribution and site-specific means well. Site-specific predictive variation was less exact: the model tended to overpredict variability at BUC3, BUL11, and LBL5 and modestly underpredict it at some other sites. We therefore retained the biologically interpretable site-specific seasonal structure rather than increasing model complexity solely to reproduce all remaining distributional features.

```{=html}
<!--
SUPPLEMENTARY MATERIAL:
The detailed model-development sequence belongs primarily in the supplement.

Suggested items:
Figure S?: Poisson vs NB vs NB + site-specific seasonal smooth comparison.
Figure S?: final density-overlay PPC.
Figure S?: PPC of mean, SD, and zeros.
Figure S?: site-specific PPC.
Figure S?: residual ACF comparison.
Table S?: MCMC diagnostics.

The main text above should remain relatively short.
-->
```

## Seasonal activity

Seasonal activity varied nonlinearly across the sampling period, and the inferred seasonal pattern changed substantially after accounting for nightly environmental conditions (@fig-seasonality). In the total seasonal model, equal-site averaged activity declined through May, increased to a first summer maximum in June, declined in July, and increased again in early August before falling through September and October (@fig-seasonality`a`). After conditioning on nightly temperature, relative humidity, and lunar phase, the seasonal trajectory was markedly different: activity was highest early in the sampling season, declined sharply from May into June, and remained comparatively low through most of the summer and autumn, with only modest increases in August and September (@fig-seasonality`b`). Thus, much of the pronounced mid-summer structure in the total seasonal pattern was associated with seasonal variation in the measured environmental conditions, while the conditional day-of-year effect was dominated by elevated activity early in the sampling season. Seasonal phenology also differed substantially among sites in both the magnitude and timing of activity (@fig-seasonality`c`), supporting the use of site-specific seasonal smooths rather than a single common seasonal trajectory.

```{r}
#| label: fig-seasonality
#| fig-cap: "Seasonal variation in community-level bat acoustic activity across eight sites in the northern Sierra Nevada, California, April–October 2015. (a) Total seasonal pattern from the day-of-year model, shown as the posterior distribution of the equal-site averaged expected number of 5-min intervals per night in which a bat call occurred. (b) Conditional seasonal pattern from the environmental model after accounting for nightly mean temperature, relative humidity, and lunar phase; the curve is likewise averaged equally across sites. (c) Site-specific seasonal phenology from the total seasonal model. Thin gray lines show posterior median trajectories for individual sites; the black line and shaded bands show the equal-site averaged seasonal trajectory. In panels a and b, and for the average in panel c, lines indicate posterior medians and shaded bands indicate 50%, 80%, and 89% credible intervals from darkest to lightest."
#| fig-width: 14
#| fig-height: 9
#| out-width: "100%"
#| echo: FALSE

include_manuscript_figure("Figure3_seasonality.PNG")
```


## Echolocation-frequency groups

Characteristic-frequency groups contributed very different amounts of total recorded activity. High-frequency species accounted for 69,001 detection intervals, compared with 27,602 for the mid-frequency group and 7,078 for the low-frequency group. Consequently, the aggregate community response was numerically dominated by high-frequency detections, although the frequency-stratified models allow the environmental responses of the three groups to be evaluated separately.

### Temperature responses among frequency groups

Temperature was positively associated with bat activity across all three frequency groups over the well-sampled portion of the observed temperature gradient, although response shape differed among groups (@fig-temperature). Across the central 80% of observed nightly temperatures (6.07–17.3 $^\circ$C), predicted activity increased from 1.52 to 13.7 intervals night$^{-1}$ for the low-frequency group, from 3.48 to 69.9 intervals night$^{-1}$ for the mid-frequency group, and from 33.9 to 68.3 intervals night${^-1}$ for the high-frequency group. The corresponding median increases were 12.2 (89% CrI: 9.50–15.5), 66.4 (54.4–81.0), and 34.3 (26.7–42.4) intervals night$^{-1}$, respectively, with posterior probability of an increase equal to 1.00 for each group. Low- and mid-frequency activity continued to increase toward the warm end of the observed gradient, whereas high-frequency activity increased through most of the well-sampled temperature range before leveling off and declining at warmer temperatures. Across posterior draws, the median temperature associated with maximum high-frequency activity was 18.8 $^\circ$C (89% CrI: 17.3–24.2 $^\circ$C); 16.9% of posterior curves placed the maximum at the upper boundary of the observed temperature range, indicating greater uncertainty in the shape of the response at the warmest temperatures.

```{r}
#| label: fig-temperature
#| fig-cap: "Conditional relationships between nightly mean temperature and bat acoustic activity for (a) low-frequency, (b) mid-frequency, and (c) high-frequency echolocation groups at eight sites in the northern Sierra Nevada, California. Predictions were generated across the observed temperature range and averaged over the observed distribution of sampled days of year, with each site weighted equally; relative humidity and lunar phase were held at their reference values. Lines show posterior medians and shaded ribbons show 50%, 80%, and 89% credible intervals (darkest to lightest). Y-axis scales differ among panels to emphasize differences in response shape among frequency groups."
#| echo: false
#| out-width: "100%"
#| fig-align: center

include_manuscript_figure("Figure4_temperature.PNG")
```

```{=html}
<!--
IMPORTANT:
Everything below this point should be populated from the FINAL frequency-group
fits, not the old manuscript models.

The old manuscript's frequency-group biological story can be used as a hypothesis
to check, NOT as a result to carry forward automatically.

Recommended order:
1. humidity — primary mechanistic prediction
2. temperature — thermal response
3. seasonal phenology
4. lunar phase — brief unless biologically notable
-->
```

### Frequency-dependent humidity responses

Relative humidity was associated with markedly different patterns of activity among frequency groups (@fig-humidity). Across the central 80% of observed nightly relative humidity conditions (49.4–84.6%), expected activity declined substantially for both the mid- and high-frequency groups. For high-frequency bats, expected activity decreased from 52.4 to 29.0 5-min detection intervals night$^{-1}$, corresponding to a median ratio of 0.55 (89% CrI: 0.50–0.62) and a posterior probability of a decrease of 1.00. Mid-frequency activity declined from 19.8 to 13.7 intervals night$^{-1}$ over the same humidity range (median ratio = 0.69, 89% CrI: 0.58–0.83; posterior probability of decrease > 0.999). In contrast, low-frequency activity showed little evidence of an overall decline, increasing slightly from 4.32 to 5.32 intervals night$^{-1}$ (median ratio = 1.23, 89% CrI: 0.97–1.56; posterior probability of decrease = 0.083). Response curves were nonlinear, particularly for the mid-frequency group, but declines became pronounced at high relative humidity for both the mid- and high-frequency groups, whereas activity of the low-frequency group remained comparatively insensitive across the observed humidity gradient.

```{r}
#| label: fig-humidity
#| fig-cap: "Conditional relationships between nightly mean relative humidity and bat activity for (a) low-, (b) mid-, and (c) high-frequency echolocation groups. Lines show posterior median expected numbers of 5-min detection intervals per night, with inner and outer ribbons representing 50% and 89% credible intervals, respectively. Predictions were generated from the frequency-group-specific Bayesian generalized additive models while holding the other environmental predictors at their reference values and averaging over the seasonal effect. Note that y-axis scales differ among panels to emphasize the shape and uncertainty of the response within each frequency group."
#| fig-width: 16
#| fig-height: 5.5
#| out-width: "100%"
#| echo: false

include_manuscript_figure("Figure5_humidity.PNG")
```


### Frequency-specific seasonal phenology

**\[PENDING FINAL FREQUENCY-GROUP FITS.\]**

```{=html}
<!--
Potentially Figure S rather than main text unless it materially explains the
aggregate seasonal decomposition.

This is where we can determine whether the residual spring/autumn pattern is
concentrated in particular frequency groups/species assemblages.
-->
```

### Lunar responses among frequency groups

Associations between lunar illumination and bat activity also differed among frequency groups (@fig-lunar). High-frequency activity increased with lunar illumination, with expected activity increasing by 10.4 5-min detection intervals per night from new moon to full moon (89% CrI: 7.02–14.4). This corresponded to a 1.28-fold increase in expected activity (89% CrI: 1.18–1.39), with essentially all posterior draws supporting an increase. Low-frequency activity showed weaker evidence of a response in the opposite direction, declining by a median of 0.92 intervals per night from new moon to full moon (89% CrI: −2.06–0.25; posterior probability of a decrease = 0.898). Mid-frequency activity exhibited a nonlinear relationship with lunar illumination, but the overall new- to full-moon contrast was comparatively small and uncertain (median difference = 1.81 intervals per night, 89% CrI: −1.00–4.80; posterior probability of an increase = 0.850). Thus, the clearest lunar response occurred among high-frequency bats, for which activity increased across the lunar illumination gradient.

```{r}
#| label: fig-lunar
#| fig-cap: "Conditional relationships between lunar illumination and bat activity for (a) low-, (b) mid-, and (c) high-frequency groups. Lunar illumination ranges from 0 (new moon) to 1 (full moon). Lines show posterior median expected numbers of 5-min detection intervals per night, with inner and outer ribbons representing 50% and 89% credible intervals, respectively. Predictions were generated from the frequency-group-specific Bayesian generalized additive models while holding temperature and relative humidity at their reference values and averaging over the seasonal effect. Note that y-axis scales differ among panels to emphasize the shape and uncertainty of the response within each frequency group."
#| fig-width: 16
#| fig-height: 5.5
#| out-width: "100%"
#| echo: false

include_manuscript_figure("Figure6_lunar.PNG")
```

## Summary of principal findings

```{=html}
<!--
WRITE THIS LAST.

For Functional Ecology, this should be a short transition into Discussion,
not a second Discussion section.

Possible structure, ONLY IF supported by final models:

"Together, the analyses separated three sources of variation that were
confounded in simpler descriptions of nightly bat activity. First, much of the
apparent seasonal trajectory reflected changing nightly environmental
conditions, while site-specific phenological structure remained after those
conditions were controlled. Second, temperature and humidity were the dominant
nightly environmental correlates of aggregate activity, whereas lunar phase had
a smaller effect. Third, humidity responses differed among characteristic
frequency groups in the direction predicted by frequency-dependent atmospheric
attenuation."

Do not activate this paragraph until all final frequency fits are complete.
-->
```

# Tables and figures to generate

```{=html}
<!--
This section is an internal production checklist. Delete it from the manuscript
before submission.

MAIN TEXT — tentative

Table 1.
Environmental and sampling summary for the corrected analysis dataset.

Figure 1.
Study map / detector locations.

Figure 2.
DAG.

Figure 3.
Total versus conditional seasonal activity.

Figure 4.
Aggregate conditional environmental effects:
temperature, RH, lunar phase.

Figure 5.
Frequency-group humidity responses.

Figure 6.
Frequency-group temperature responses, if biologically informative enough for
main text. Otherwise move to supplement.

SUPPLEMENT — tentative

Table S1.
Model definitions / estimands.

Table S2.
Aggregate-model diagnostics.

Table S3.
Frequency-group model diagnostics.

Table S4.
Species, characteristic frequency, frequency group, detection totals.

Figure S1.
Site-specific seasonal smooths.

Figure S2.
Model-development ACF comparison.

Figure S3.
Aggregate PPCs.

Figure S4.
Site-specific PPCs.

Figure S5+.
Frequency-group PPCs / residual ACFs.

Additional sensitivity figure:
High-frequency temperature response with and without low-activity/cold sites,
ONLY if still scientifically necessary after final model fitting.
-->
```
