# Results

```{r}
#| label: results-setup
#| include: false

library(knitr)

# Project root
project_root <- normalizePath(
  file.path(getwd(), ".."),
  winslash = "/",
  mustWork = TRUE
)

# Manuscript figure directory
fig_dir <- file.path(project_root, "figs")

# Derived manuscript table directory
table_dir <- file.path(
  project_root,
  "derived",
  "tables"
)

# Helper for manuscript figures
fig_path <- function(filename) {
  file.path(fig_dir, filename)
}

# Helper for derived manuscript tables
table_path <- function(filename) {
  file.path(table_dir, filename)
}

# Fail clearly if a required manuscript file is missing
require_file <- function(path) {
  if (!file.exists(path)) {
    stop("Required manuscript file not found: ", path)
  }
  invisible(path)
}

# Include an already-generated manuscript figure
include_manuscript_figure <- function(filename) {
  path <- fig_path(filename)
  require_file(path)

  knitr::include_graphics(
    path,
    rel_path = FALSE
  )
}

```

## Sampling and acoustic activity

The analysis comprised 1,056 sampled site-nights across eight sites and 179 sampling nights from late April through mid-October 2015 (@tbl-study-summary). Nights without usable acoustic recordings were excluded as unsampled rather than treated as zero activity; sampled nights with no accepted bat detections were retained as legitimate zeroes. Sampling effort therefore varied among sites because of detector failures and other periods without usable acoustic data, ranging from 47 retained nights at BUL18 to 174 nights at the most intensively sampled sites.

Across the retained site-nights, aggregate bat activity totaled 103,681 5-min detection intervals representing 17 species. Of these detections, 7,078 were assigned to the low-frequency group, 27,602 to the mid-frequency group, and 69,001 to the high-frequency group. Environmental conditions represented in the analysis spanned substantial variation in nightly temperature, relative humidity, and lunar phase (@tbl-study-summary).

```{r}
#| label: tbl-study-summary
#| tbl-cap: "Summary of the acoustic sampling design and environmental conditions represented in the analysis. Environmental ranges are based on the site-night observations retained for modeling."
#| echo: false

table_1_file <- table_path("table_1.rds")
require_file(table_1_file)

table_1 <- readRDS(table_1_file)

knitr::kable(
  table_1,
  align = c("l", "c")
)
```

## Model adequacy and diagnostics

The models represented site-specific seasonal phenology using penalized factor-smooth interactions between day of year and site. In the aggregate conditional model, residual autocorrelation declined from 0.379 at a one-day lag to 0.208 at two days and 0.168 at three days, approached zero by four days, and remained weak at longer lags. Remaining temporal dependence was therefore concentrated primarily at the shortest lags.

MCMC diagnostics indicated satisfactory posterior sampling. All reported $\hat{R}$ values were 1.00, with large bulk and tail effective sample sizes, and seven of 12,000 post-warmup transitions in the aggregate conditional model reached the maximum tree depth of 12. Posterior predictive checks reproduced the overall distribution of nightly activity and site-specific mean activity well. Predictive variance was overestimated at BUC3, BUL11, and LBL5 and modestly underestimated at several other sites, but the principal seasonal, among-site, and distributional features relevant to the environmental contrasts were adequately represented. Detailed sampling diagnostics, posterior predictive checks, and residual temporal diagnostics are provided in the Supplementary Material.

## Seasonal activity

Seasonal activity varied nonlinearly across the sampling period, and the inferred seasonal pattern changed substantially after accounting for nightly environmental conditions (@fig-seasonality). In the total seasonal model, activity averaged equally across sites declined through May, increased to a first summer maximum in June, declined in July, and increased again in early August before falling through September and October (@fig-seasonality`a`). After conditioning on nightly temperature, relative humidity, and lunar phase, the seasonal trajectory was markedly different: activity was highest early in the sampling season, declined sharply from May into June, and remained comparatively low through most of the summer and autumn, with only modest increases in August and September (@fig-seasonality`b`). Thus, much of the pronounced mid-summer structure in the total seasonal pattern was associated with seasonal variation in the measured environmental conditions, while the conditional day-of-year effect was dominated by elevated activity early in the sampling season. Seasonal phenology also differed substantially among sites in both the magnitude and timing of activity (@fig-seasonality`c`), supporting the use of site-specific seasonal smooths rather than a single common seasonal trajectory.

```{r}
#| label: fig-seasonality
#| fig-cap: "Seasonal variation in community-level bat acoustic activity across eight sites in the northern Sierra Nevada, California, April–October 2015. (a) Total seasonal pattern from the day-of-year model, shown as the posterior distribution of the equal-site averaged expected number of 5-min intervals per night in which a bat call occurred. (b) Conditional seasonal pattern from the environmental model after accounting for nightly mean temperature, relative humidity, and lunar phase; the curve is likewise averaged equally across sites. (c) Site-specific seasonal phenology from the total seasonal model. Thin gray lines show posterior median trajectories for individual sites; the black line and shaded bands show the equal-site averaged seasonal trajectory. In panels a and b, and for the average in panel c, lines indicate posterior medians and shaded bands indicate 50%, 80%, and 89% credible intervals from darkest to lightest."
#| fig-width: 14
#| fig-height: 9
#| out-width: "100%"
#| echo: FALSE

include_manuscript_figure("Figure3_seasonality.PNG")
```


## Echolocation-frequency groups

Characteristic-frequency groups contributed very different amounts of total recorded activity. High-frequency species accounted for 69,001 detection intervals, compared with 27,602 for the mid-frequency group and 7,078 for the low-frequency group. Consequently, the aggregate community response was numerically dominated by high-frequency detections, although the frequency-stratified models allow the environmental responses of the three groups to be evaluated separately.


### Frequency-dependent humidity responses

Relative humidity was associated with systematically different patterns of acoustic activity among frequency groups (@fig-humidity). Across the central 80% of observed nightly relative humidity conditions (49.4–84.6%), expected activity declined substantially for both the mid- and high-frequency groups, with the strongest decline occurring in the high-frequency group. For high-frequency bats, expected activity decreased from 52.4 to 29.0 5-min detection intervals per night, corresponding to a median ratio of 0.55 (89% CrI: 0.50–0.62) and a posterior probability of a decrease of 1.00. Mid-frequency activity declined from 19.8 to 13.7 intervals per night over the same humidity range (median ratio = 0.69, 89% CrI: 0.58–0.83; posterior probability of decrease > 0.999). In contrast, low-frequency activity showed little evidence of an overall decline, increasing slightly from 4.32 to 5.32 intervals per night (median ratio = 1.23, 89% CrI: 0.97–1.56; posterior probability of decrease = 0.083). Response curves were nonlinear, particularly for the mid-frequency group, but declines became pronounced at high relative humidity for both the mid- and high-frequency groups, whereas activity of the low-frequency group remained comparatively insensitive across the observed humidity gradient.

```{r}
#| label: fig-humidity
#| fig-cap: "Conditional relationships between nightly mean relative humidity and bat acoustic activity for (a) low-, (b) mid-, and (c) high-frequency echolocation groups. Lines show posterior median expected numbers of 5-min detection intervals per night, with inner and outer ribbons representing 50% and 89% credible intervals, respectively. Predictions were generated across the observed relative-humidity range and averaged over the observed distribution of sampled days of year, with each site weighted equally; temperature and lunar phase were held at their reference values. Y-axis scales differ among panels to emphasize response shape and uncertainty within each frequency group."
#| fig-width: 16
#| fig-height: 5.5
#| out-width: "100%"
#| echo: false

include_manuscript_figure("Figure4_humidity.PNG")
```

### Temperature responses among frequency groups

Temperature was positively associated with bat activity across all three frequency groups over the well-sampled portion of the observed temperature gradient, although response shape differed among groups (@fig-temperature). Across the central 80% of observed nightly temperatures (6.07–17.3 $^\circ$C), predicted activity increased from 1.52 to 13.7 intervals per night for the low-frequency group, from 3.48 to 69.9 intervals per night for the mid-frequency group, and from 33.9 to 68.3 intervals per night for the high-frequency group. The corresponding median increases were 12.2 (89% CrI: 9.50–15.5), 66.4 (54.4–81.0), and 34.3 (26.7–42.4) intervals per night, respectively, with posterior probability of an increase equal to 1.00 for each group. Low- and mid-frequency activity continued to increase toward the warm end of the observed gradient, whereas high-frequency activity increased through most of the well-sampled temperature range before leveling off and declining at warmer temperatures. Across posterior draws, the median temperature associated with maximum high-frequency activity was 18.8 $^\circ$C (89% CrI: 17.3–24.2 $^\circ$C); 16.9% of posterior curves placed the maximum at the upper boundary of the observed temperature range, indicating greater uncertainty in the shape of the response at the warmest temperatures.

```{r}
#| label: fig-temperature
#| fig-cap: "Conditional relationships between nightly mean temperature and bat acoustic activity for (a) low-frequency, (b) mid-frequency, and (c) high-frequency echolocation groups at eight sites in the northern Sierra Nevada, California. Predictions were generated across the observed temperature range and averaged over the observed distribution of sampled days of year, with each site weighted equally; relative humidity and lunar phase were held at their reference values. Lines show posterior medians and shaded ribbons show 50%, 80%, and 89% credible intervals (darkest to lightest). Y-axis scales differ among panels to emphasize differences in response shape among frequency groups."
#| echo: false
#| out-width: "100%"
#| fig-align: center

include_manuscript_figure("Figure5_temperature.PNG")
```

### Lunar-phase responses among frequency groups

Associations between lunar phase and bat acoustic activity differed among frequency groups (@fig-lunar). High-frequency activity increased from new moon toward full moon, with expected activity increasing by 10.4 5-min detection intervals per night between the two endpoints (89% CrI: 7.02–14.4). This corresponded to a 1.28-fold increase in expected activity (89% CrI: 1.18–1.39), with essentially all posterior draws supporting an increase. Low-frequency activity showed weaker evidence of a response in the opposite direction, declining by a median of 0.92 intervals per night from new moon to full moon (89% CrI: −2.06–0.25; posterior probability of a decrease = 0.898). Mid-frequency activity exhibited a nonlinear relationship with lunar phase, but the overall new- to full-moon contrast was comparatively small and uncertain (median difference = 1.81 intervals per night, 89% CrI: −1.00–4.80; posterior probability of an increase = 0.850). Thus, the clearest lunar-phase response occurred among high-frequency bats, for which activity was higher toward the full-moon end of the lunar cycle.

```{r}
#| label: fig-lunar
#| fig-cap: "Conditional relationships between lunar phase and bat acoustic activity for (a) low-, (b) mid-, and (c) high-frequency echolocation groups. Lunar phase is represented from 0 (new moon) to 1 (full moon). Lines show posterior median expected numbers of 5-min detection intervals per night, with inner and outer ribbons representing 50% and 89% credible intervals, respectively. Predictions were generated across the observed lunar-phase range and averaged over the observed distribution of sampled days of year, with each site weighted equally; temperature and relative humidity were held at their reference values. Y-axis scales differ among panels to emphasize response shape and uncertainty within each frequency group."
#| fig-width: 16
#| fig-height: 5.5
#| out-width: "100%"
#| echo: false

include_manuscript_figure("Figure6_lunar.PNG")
```

### Environmental contrasts

Posterior contrasts summarized the differing environmental responses among frequency groups (@tbl-environmental-contrasts). Increasing temperature from the 10th to the 90th percentile of observed nightly conditions was associated with higher predicted activity in all three groups, although the proportional increase was substantially larger for the low- and mid-frequency groups than for the high-frequency group. In contrast, increasing relative humidity over the same portion of its observed distribution was associated with pronounced declines in mid- and high-frequency activity, whereas the low-frequency response was comparatively weak. Lunar phase was associated with smaller and less consistent differences across groups, except for the high-frequency group, for which predicted activity was higher under full- than new-moon conditions.

```{=openxml}
<w:p>
  <w:pPr>
    <w:sectPr>
      <w:type w:val="nextPage"/>
      <w:pgSz w:w="12240" w:h="15840"/>
    </w:sectPr>
  </w:pPr>
</w:p>
```

```{r}
#| label: tbl-environmental-contrasts
#| tbl-cap: "Posterior contrasts in expected nightly bat activity across environmental gradients for low-, mid-, and high-frequency echolocation groups. Temperature and relative-humidity contrasts compare predictions at the 10th and 90th percentiles of the observed covariate distributions; lunar contrasts compare new moon (0) with full moon (1). Differences and ratios are posterior medians with 89% credible intervals. P(Δ > 0) is the posterior probability that predicted activity at the upper endpoint exceeds predicted activity at the lower endpoint. Predictions are conditional on the remaining covariates in the full environmental model."
#| echo: false

table_2_file <- table_path("table_2.rds")
require_file(table_2_file)

table_2 <- readRDS(table_2_file)

knitr::kable(
  table_2,
  align = c("l", "l", "l", "r", "r", "c", "c", "c")
)
```

```{=openxml}
<w:p>
  <w:pPr>
    <w:sectPr>
      <w:type w:val="nextPage"/>
      <w:pgSz w:w="15840" w:h="12240" w:orient="landscape"/>
    </w:sectPr>
  </w:pPr>
</w:p>
```
