# Supplementary material {.unnumbered}

This supplementary material contains model definitions, sampling and species summaries, model diagnostics, posterior predictive checks, residual temporal diagnostics, and sensitivity analyses supporting the main-text analysis of bat activity in the northern Sierra Nevada. Intermediate model-development objects that do not contribute directly to interpretation are retained in the analysis repository rather than reproduced here.

```{r}
#| label: setup-supplement
#| include: false

library(tidyverse)
library(knitr)

table_dir  <- "tables"
figure_dir <- "figures"
derived_dir <- "derived"

# If this file lives in manuscript/, change these to ../tables, ../figures, etc.
```

# Supplementary tables

## Table S1. Model definitions and estimands

Table S1 documents the final model set, response distribution, seasonal structure, environmental covariates, and interpretation of each model.

```{r}
#| label: tbl-s1-model-definitions
#| tbl-cap: "Final Bayesian generalized additive models used to estimate aggregate and frequency-group bat activity. The table distinguishes the total seasonal estimand from conditional environmental associations."
#| echo: false
#| warning: false
#| message: false

model_def_file <- file.path(table_dir, "model_definitions.csv")

if (file.exists(model_def_file)) {
  model_definitions <- readr::read_csv(
    model_def_file,
    show_col_types = FALSE
  )

  knitr::kable(model_definitions)
} else {
  cat("PLACEHOLDER: generate tables/model_definitions.csv from the final model lookup.")
}
```

```{=html}
<!--
Suggested columns:
Dataset
Model
Response
Family
Formula / seasonal structure
Environmental covariates
Primary estimand

At minimum include:
- Aggregate season-only
- Aggregate conditional environmental
- Low-frequency conditional environmental
- Mid-frequency conditional environmental
- High-frequency conditional environmental

Add frequency-group season-only models only if they are retained for inference.
-->
```

## Table S2. Aggregate-model sampling diagnostics

```{r}
#| label: tbl-s2-aggregate-diagnostics
#| tbl-cap: "Sampling diagnostics for the final aggregate Bayesian generalized additive models."
#| echo: false
#| warning: false
#| message: false

diag_file <- file.path(table_dir, "aggregate_model_diagnostics.csv")

if (file.exists(diag_file)) {
  aggregate_diagnostics <- readr::read_csv(
    diag_file,
    show_col_types = FALSE
  )

  knitr::kable(aggregate_diagnostics)
} else {
  cat("PLACEHOLDER: generate aggregate_model_diagnostics.csv.")
}
```

```{=html}
<!--
Suggested columns:
Model
N site-nights
Chains
Post-warmup draws
Max R-hat
Min bulk ESS
Min tail ESS
Divergences
Maximum-treedepth transitions
E-BFMI warnings
Negative-binomial shape (median, interval)
-->
```

## Table S3. Frequency-group model sampling diagnostics

```{r}
#| label: tbl-s3-frequency-diagnostics
#| tbl-cap: "Sampling diagnostics for the final low-, mid-, and high-frequency Bayesian generalized additive models."
#| echo: false
#| warning: false
#| message: false

diag_file <- file.path(table_dir, "frequency_model_diagnostics.csv")

if (file.exists(diag_file)) {
  frequency_diagnostics <- readr::read_csv(
    diag_file,
    show_col_types = FALSE
  )

  knitr::kable(frequency_diagnostics)
} else {
  cat("PLACEHOLDER: populate after the frequency-group production fits are finalized.")
}
```

## Table S4. Species composition and characteristic echolocation-frequency groups

The low-frequency group contains *Lasiurus cinereus*, *Eumops perotis*, and *Euderma maculatum*. The mid-frequency group contains *Eptesicus fuscus*, *Lasionycteris noctivagans*, *Tadarida brasiliensis*, *Antrozous pallidus*, *Myotis evotis*, *Myotis thysanodes*, and *Corynorhinus townsendii*. The high-frequency group contains *Myotis yumanensis*, *M. lucifugus*, *M. californicus*, *M. ciliolabrum*, *M. volans*, *Parastrellus hesperus*, and *Lasiurus frantzii*.

```{r}
#| label: tbl-s4-species
#| tbl-cap: "Bat species included in the analysis, characteristic echolocation frequency, assigned frequency group, and acoustic detection summary."
#| echo: false
#| warning: false
#| message: false

species_file <- file.path(table_dir, "species_frequency_summary.csv")

if (file.exists(species_file)) {
  species_summary <- readr::read_csv(
    species_file,
    show_col_types = FALSE
  )

  knitr::kable(species_summary)
} else {
  cat("PLACEHOLDER: generate species_frequency_summary.csv from species_freq and the corrected analytical data.")
}
```

## Table S5. Detector effort and retained site-nights

```{r}
#| label: tbl-s5-detector-effort
#| tbl-cap: "Acoustic sampling effort by site after distinguishing detector-failure nights from sampled nights with zero bat activity."
#| echo: false
#| warning: false
#| message: false

effort_file <- file.path(table_dir, "detector_effort_summary.csv")

if (file.exists(effort_file)) {
  effort_summary <- readr::read_csv(
    effort_file,
    show_col_types = FALSE
  )

  knitr::kable(effort_summary)
} else {
  cat("PLACEHOLDER: generate detector_effort_summary.csv.")
}
```

```{=html}
<!--
Expected retained site-nights:
BUC3 172
BUL11 174
BUL18 47
BUL4 64
GRP1 174
GRP5 168
LBL5 174
THL1 83
Total = 1,056
-->
```

## Table S6. Software and computational environment

```{r}
#| label: tbl-s6-software
#| tbl-cap: "Software and package versions used in the final analysis."
#| echo: false
#| warning: false
#| message: false

software_file <- file.path(table_dir, "software_versions.csv")

if (file.exists(software_file)) {
  software_versions <- readr::read_csv(
    software_file,
    show_col_types = FALSE
  )

  knitr::kable(software_versions)
} else {
  cat("PLACEHOLDER: run write_package_bib() to generate software_versions.csv.")
}
```

# Supplementary figures

## Figure S1. Site-specific seasonal phenology

Site-specific seasonal smooths from the final aggregate model illustrate how the timing and magnitude of activity varied among sites around the shared seasonal trajectory.

```{r}
#| label: fig-s1-site-phenology
#| fig-cap: "Site-specific seasonal phenology in bat activity from the final aggregate model. Curves show posterior predictions across day of year for each site under the fitted factor-smooth structure. Lines represent posterior medians and ribbons show 89% credible intervals."
#| echo: false
#| warning: false
#| message: false

fig_file <- file.path(figure_dir, "site_specific_seasonal_phenology.png")

if (file.exists(fig_file)) {
  knitr::include_graphics(fig_file)
} else {
  cat("PLACEHOLDER: generate site-specific seasonal phenology figure.")
}
```

## Figure S2. Residual temporal dependence during model development

This figure documents how residual temporal structure changed as the likelihood and seasonal structure were revised. Calendar-day lags are used so gaps in sampling are not treated as consecutive observations.

```{r}
#| label: fig-s2-acf-comparison
#| fig-cap: "Calendar-aware residual autocorrelation for candidate aggregate models. The comparison illustrates the persistence of temporal dependence in simpler models and its reduction after site-specific seasonal phenology was represented."
#| echo: false
#| warning: false
#| message: false

fig_file <- file.path(figure_dir, "aggregate_acf_model_comparison.png")

if (file.exists(fig_file)) {
  knitr::include_graphics(fig_file)
} else {
  cat("PLACEHOLDER: generate aggregate ACF model-comparison figure.")
}
```

## Figure S3. Aggregate posterior predictive checks

```{r}
#| label: fig-s3-aggregate-ppc
#| fig-cap: "Posterior predictive checks for the final aggregate conditional environmental model, including the overall response distribution and test statistics for the mean, standard deviation, and frequency of zero-activity nights."
#| echo: false
#| warning: false
#| message: false

fig_file <- file.path(figure_dir, "aggregate_ppc.png")

if (file.exists(fig_file)) {
  knitr::include_graphics(fig_file)
} else {
  cat("PLACEHOLDER: assemble aggregate posterior predictive checks.")
}
```

## Figure S4. Site-specific posterior predictive checks

```{r}
#| label: fig-s4-site-ppc
#| fig-cap: "Observed and posterior-predictive summaries of nightly bat activity by site for the final aggregate conditional environmental model."
#| echo: false
#| warning: false
#| message: false

fig_file <- file.path(figure_dir, "site_specific_ppc.png")

if (file.exists(fig_file)) {
  knitr::include_graphics(fig_file)
} else {
  cat("PLACEHOLDER: generate site-specific posterior predictive checks.")
}
```

## Figure S5. Frequency-group posterior predictive checks and residual autocorrelation

```{r}
#| label: fig-s5-frequency-diagnostics
#| fig-cap: "Posterior predictive checks and calendar-aware residual autocorrelation for the final low-, mid-, and high-frequency bat activity models."
#| echo: false
#| warning: false
#| message: false

fig_file <- file.path(figure_dir, "frequency_group_diagnostics.png")

if (file.exists(fig_file)) {
  knitr::include_graphics(fig_file)
} else {
  cat("PLACEHOLDER: populate after final frequency-group models are accepted.")
}
```

## Figure S6. High-frequency temperature sensitivity analysis

```{=html}
<!--
The previous manuscript used Figure S6 to compare the high-frequency
temperature response with and without BUL4, BUL18, and THL1. Retain this figure
ONLY if the unusual cool-temperature response persists under the corrected
sampling frame and site-specific seasonal factor smooth.

If it no longer persists or the revised model makes the site-filtering analysis
unnecessary, delete this section rather than carrying the old sensitivity
analysis forward.
-->
```

```{r}
#| label: fig-s6-high-temp-sensitivity
#| fig-cap: "Sensitivity analysis for the high-frequency temperature response. This figure will be retained only if a cool-temperature peak persists in the final frequency-group model and requires evaluation of site-specific support."
#| echo: false
#| warning: false
#| message: false

fig_file <- file.path(figure_dir, "high_frequency_temperature_sensitivity.png")

if (file.exists(fig_file)) {
  knitr::include_graphics(fig_file)
} else {
  cat("PLACEHOLDER: conditional on the final high-frequency model.")
}
```

# Additional model assessment

Routine trace plots need not be reproduced if the tabulated sampling diagnostics adequately document convergence and sampling quality. Additional graphical diagnostics should be retained when they evaluate a specific concern raised during model development, particularly posterior predictive fit, residual temporal structure, or sensitivity of a biological conclusion to site composition.

```{=html}
<!--
SUPPLEMENT UPDATE CHECKLIST

1. Generate Tables S1-S6 directly from pipeline outputs.
2. Update all figure filenames to match figure_functions.R.
3. Decide whether frequency-group seasonal curves belong in main text or supplement.
4. Decide whether lunar group curves warrant a supplementary figure.
5. Retain Figure S6 only if the high-frequency cold-temperature pattern persists.
6. Add any model-development comparison requested by reviewers, but avoid dumping
   every exploratory diagnostic into the supplement.
7. All numerical diagnostics must come from FINAL four-chain fits, not one-chain
   diagnostic models.
-->
```
