# Supplementary material {.unnumbered}

This supplementary material provides model specifications, sampling and species summaries, model diagnostics, posterior predictive checks, and residual temporal diagnostics supporting the main-text analysis of bat activity in the northern Sierra Nevada.

```{r}
#| label: setup-supplement
#| include: false

library(tidyverse)
library(knitr)

table_dir  <- file.path("..", "derived", "tables")
figure_dir <- normalizePath(
  "../figs/supplementary",
  winslash = "/",
  mustWork = TRUE
)

require_file <- function(path) {
  if (!file.exists(path)) {
    stop(
      "Required supplementary output not found: ",
      path,
      call. = FALSE
    )
  }
  invisible(path)
}

supp_table <- function(name) {
  path <- file.path(
    table_dir,
    paste0(name, ".rds")
  )

  require_file(path)
  readRDS(path)
}

include_supp_figure <- function(filename) {

  path <- file.path(
    figure_dir,
    filename
  )

  require_file(path)

  knitr::include_graphics(
    path,
    rel_path = FALSE
  )
}

```

# Supplementary tables

## Table S1. Final model specifications

```{r}
#| label: tbl-s1-model-definitions
#| tbl-cap: "Final Bayesian generalized additive models used to estimate aggregate and frequency-group bat activity. All models used a negative-binomial response distribution with a log link and site-specific seasonal phenology represented by penalized factor-smooth interactions where applicable."
#| echo: false
#| warning: false
#| message: false

table_s1 <- supp_table("table_s1")

knitr::kable(
  table_s1,
  align = c("l", "l", "l", "l", "r")
)
```

## Table S2. Aggregate-model sampling diagnostics

```{r}
#| label: tbl-s2-aggregate-diagnostics
#| tbl-cap: "MCMC sampling diagnostics for the final aggregate seasonal and conditional environmental models. Negative-binomial shape is summarized by the posterior median and 89% credible interval."
#| echo: false
#| warning: false
#| message: false

table_s2 <- supp_table("table_s2")

knitr::kable(
  table_s2,
  digits = c(NA, 0, 0, 2, 0, 0, 0, 0, 2, 2, 2)
)
```

## Table S3. Frequency-group model sampling diagnostics

```{r}
#| label: tbl-s3-frequency-diagnostics
#| tbl-cap: "MCMC sampling diagnostics for the final low-, mid-, and high-frequency conditional environmental models. Negative-binomial shape is summarized by the posterior median and 89% credible interval."
#| echo: false
#| warning: false
#| message: false

table_s3 <- supp_table("table_s3")

knitr::kable(
  table_s3,
  digits = c(NA, 0, 0, 2, 0, 0, 0, 0, 2, 2, 2)
)
```

## Table S4. Species composition and characteristic echolocation-frequency groups

```{r}
#| label: tbl-s4-species
#| tbl-cap: "Bat species included in the analysis, characteristic echolocation frequency, assigned frequency group, total number of 5-min detection intervals, and number of retained site-nights on which each species was detected."
#| echo: false
#| warning: false
#| message: false

table_s4 <- supp_table("table_s4")

knitr::kable(
  table_s4,
  align = c("l", "l", "l", "r", "l", "r", "r")
)
```

## Table S5. Retained sampling effort by site

```{r}
#| label: tbl-s5-detector-effort
#| tbl-cap: "Retained acoustic sampling effort and activity summaries by site after nights without usable acoustic recordings were excluded. Zero-activity nights are sampled nights with usable recordings but no accepted bat detections."
#| echo: false
#| warning: false
#| message: false

table_s5 <- supp_table("table_s5")

knitr::kable(
  table_s5,
  digits = c(NA, 0, 0, 0, 1, 1, NA, NA)
)
```

## Table S6. Software and computational environment

```{r}
#| label: tbl-s6-software
#| tbl-cap: "Software and package versions used in the final analysis."
#| echo: false
#| warning: false
#| message: false

table_s6 <- supp_table("table_s6")

knitr::kable(table_s6)
```

# Supplementary figures

## Figure S1. Site-specific seasonal phenology

```{r}
#| label: fig-s1-site-phenology
#| fig-cap: "Site-specific seasonal phenology in bat activity from the final aggregate seasonal model. Gray curves show posterior median activity trajectories for individual sites, and the black curve and shaded intervals show the equal-site posterior average and associated credible intervals."
#| echo: false
#| warning: false
#| message: false
#| out-width: "100%"

include_supp_figure(
  "Figure_S1_site_seasonality.PNG"
)

```

## Figure S2. Residual temporal dependence

```{r}
#| label: fig-s2-acf-comparison
#| fig-cap: "Calendar-aware residual correlation for the final aggregate conditional environmental model and the final low-, mid-, and high-frequency conditional environmental models. Correlations are calculated using actual calendar-day lags so that gaps in sampling are not treated as consecutive observations."
#| echo: false
#| warning: false
#| message: false
#| out-width: "100%"

include_supp_figure("Figure_S2_residual_acf.PNG")

```

## Figure S3. Aggregate posterior predictive checks

```{r}
#| label: fig-s3-aggregate-ppc
#| fig-cap: "Posterior predictive checks for the final aggregate conditional environmental model. Checks compare the observed nightly activity distribution with replicated datasets generated from the posterior predictive distribution and evaluate key features of the response distribution."
#| echo: false
#| warning: false
#| message: false
#| out-width: "100%"

include_supp_figure("Figure_S3_aggregate_ppc.PNG")

```

## Figure S4. Site-specific posterior predictive checks

```{r}
#| label: fig-s4-site-ppc
#| fig-cap: "Site-specific posterior predictive checks for the final aggregate conditional environmental model. Crosses show observed site-level means and standard deviations of nightly activity; circles show posterior predictive medians and vertical intervals show 89% posterior predictive intervals."
#| echo: false
#| warning: false
#| message: false
#| out-width: "100%"

include_supp_figure("Figure_S4_site_ppc.PNG")

```

## Figure S5. Frequency-group posterior predictive checks

```{r}
#| label: fig-s5-frequency-diagnostics
#| fig-cap: "Posterior predictive checks for the final low-, mid-, and high-frequency conditional environmental models. For each frequency group, checks compare the observed response distribution with posterior predictive replicates and evaluate the standard deviation and proportion of zero-activity nights."
#| echo: false
#| warning: false
#| message: false
#| out-width: "100%"

include_supp_figure("Figure_S5_frequency_ppc.PNG")

```

# Additional model assessment

Routine trace plots are not reproduced because the tabulated sampling diagnostics document convergence and effective sample sizes for the final models. The graphical diagnostics above focus on model adequacy relevant to the biological estimands: posterior predictive fit, among-site predictive performance, and remaining short-lag temporal dependence.
